---
title: "Basic Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidybert)
library(torch)
library(torchtransformers)
library(dplyr)
library(tibble)
library(stringr)
library(tidyr)
library(dlr)
library(rsample)
```

The tidybert package, along with {[torchtransformers](https://github.com/macmillancontentscience/torchtransformers)}, provides an R interface to the BERT family of language models.^[It is the {[torch](https://torch.mlverse.org/)}-based successor to the TensorFlow-based package {[RBERT](https://github.com/jonathanbratt/RBERT)}.]

Generally speaking, there are three levels at which a BERT model could be used:

1. Using the output embeddings produced by a pre-trained BERT model as features in a task-specific model.
2. Fine-tuning a task-specific deep learning model that includes pre-trained BERT weights as initial values that are subsequently refined.
3. Training a BERT model from scratch (and then possibly using the pre-trained model in one of the two ways above).


{tidybert} exports functions to make 1. and 2. as easy as possible. 
It also enables exploring the output embeddings and attention weights produced by a BERT model, to better understand how these models work.

# Producing contextual embeddings

Assume we have a data frame of input text. 
Each input sequence is composed of two segments, with each segment in a different column in the table.

```{r}
text_inputs <- tibble::tribble(
  ~segment1, ~segment2,
  "Why didn't the chicken cross the road?", "Because it was too wide.",
  "Why didn't the chicken cross the road?", "Because it was too tired.",
  "Why didn't the chicken cross the road?", "Because it was raining."
)
text_inputs
```

We want to produce contextual embeddings for these sequences.
First we tokenize the input text into the format required by the model. 
We'll use the `tokenize_bert` function exported from the {torchtransformers} package (the api here will likely be
changing soon).

```{r}
# Set the number of tokens to pad/truncate the result to.
n_tokens <- 20L
n_inputs <- length(text_inputs)
tokenized <- torchtransformers::tokenize_bert(text_inputs$segment1,
                                              text_inputs$segment2,
                                              n_tokens = n_tokens)
```

Load a pretrained BERT model and pass in the input:

```{r}
bert_model <- torchtransformers::make_and_load_bert(
  model_name = "bert_base_uncased"
)

bert_model$eval() 
bert_output <- bert_model(torch::torch_tensor(tokenized$token_ids), 
                          torch::torch_tensor(tokenized$token_type_ids))
```

## Extracting embeddings in tidy form

Everything up to this point was done with the {torchtransformers} package. 
The {tidybert} package provides functions to further process the model output into a tidy format.^[The tidied outputs are in a format that is compatible with the visualization functions from the package {[RBERTviz](https://github.com/jonathanbratt/RBERTviz)}.]

```{r}
tidy_out <- tidybert::tidy_bert_output(bert_output, tokenized)
```

The `embeddings` element of `tidy_out` is a data frame containing one embedding vector per row. 
Each row also contains columns identifying the input sequence, segment, token, and layer. 

Note that we use 1-based indexing, so a 12-layer BERT model (for example) is indexed from 1 to 12. 
The "0" layer in this table corresponds to the base embeddings *before* layer 1, that is, before any layers of self-attention.

```{r}
# glimpse just the first few columns to understand the structure
dplyr::glimpse(tidy_out$embeddings[, 1:10])
```


## Extracting the attention weights

The tidied output also contains a data frame with the attention weights.
Each row contains the weight value for the attention paid from a single token to a single token, along with columns identifying the sequence, the tokens, the segments, the layer, and the attention "head".

```{r}
dplyr::glimpse(tidy_out$attention)
```


# Fine-tuning models

{tidybert} also provides functions to simplify the fine-tuning of BERT-based classification models. 
(A simple classifier head consisting of a dense layer is used; for more control, build a custom model directly with {torchtransformers}.)

To illustrate the process, let's make a classifier for synonyms/antonyms.
This is an interesting example case, because one of the presumed benefits to using word vectors is that the embeddings provide a measure of the similarity between words. 
So, it _should_ be easy to train a model to distinguish between synonyms and antonyms.

First, let's get some training data...

```{r, eval = FALSE}
# Set up a processor function for {dlr} to load the data.
process_synant <- function(source_file) {
  synant <- readr::read_csv(source_file)

  # This is an extremely messy table. We're not going to try to do any real
  # cleaning, just throw away the messy parts. 
  # Fix the names:
  names(synant) <- c("word", "synonym", "antonym")
  
  # The first 103 rows are mixed with text, so let's throw those away.
  synant_clean <- synant %>% 
    dplyr::filter(dplyr::row_number() > 103) %>% 
    # Throw away any rows that have "[see also]" notes:
    dplyr::filter(!grepl("[", synonym, fixed = TRUE),
                  !grepl("[", antonym, fixed = TRUE),
                  !grepl("[", word, fixed = TRUE),
                  # Some rows have errors that probably came from formatting
                  !grepl("SYN", word, fixed = TRUE),
                  # The easiest way to remove extra end matter is...
                  !is.na(synonym),
                  !is.na(antonym)
    ) %>% 
    # sometimes the word can be found verbatim in the synonym list. This looks
    # like it's usually a formatting error, but we want to avoid that anyway.
    # Toss 'em!
    dplyr::rowwise() %>% 
    dplyr::filter(!grepl(word, synonym, fixed = TRUE))
  # # Some words have part-of-speech flags. Keep those for this.
  
  # One more thing: let's break up the lists so there's a single antonym or
  # synonym per example. A nice feature of this data is that double-spaces are
  # used to delineate separate entries, so multi-word phrases can still be
  # preserved.
  # Let's also pivot so that the synonyms and antonyms are in the same column.
  synant_tidy <- synant_clean %>% 
    tidyr::pivot_longer(cols = dplyr::all_of(c("antonym", 
                                               "synonym")),
                        names_to = "target",
                        values_to = "other_word") %>% 
    dplyr::mutate(target = as.factor(target)) %>% 
    dplyr::mutate(other_word = stringr::str_split(other_word, "\\s{2,5}")) %>% 
    tidyr::unnest_longer(other_word)
  return(synant_tidy)
}

synant_url <- "https://www.gutenberg.org/files/51155/old/20190304-51155.csv"
synant_tidy <- dlr::read_or_cache(
  source_path = synant_url,
  appname = "torchtransformers",
  process_f = process_synant
)


table(synant_tidy$target)
# antonym synonym 
#   31411   39552 
```

Wow! We now have a data set of over 71K examples of antonyms and synonyms.

```{r, eval = FALSE}
# Let's split off some rows for testing...
train_test_split <- rsample::initial_split(synant_tidy,
                                           prop = 0.9,
                                           strata = target)

training_set <- rsample::training(train_test_split)
testing_set <- rsample::testing(train_test_split)
```

Now build a classification model.
We are using two-segment input sequences, with the word as one segment and the other word as the other sequence. 
The minimum number of tokens we will need is five, but some of the words will comprise multiple tokens, so we'll give ourselves some extra space.

```{r, eval = FALSE}
# This runs in less than 20 minutes on a laptop CPU.
bert_classifier <- bert_classification(
  x = dplyr::select(training_set, word, other_word),
  y = dplyr::select(training_set, target),
  model_name = "bert_tiny_uncased",
  n_tokens = 14L, # don't need many tokens for this
  epochs = 5L
)
```

Let's see what the predictions are on the test set.

```{r, eval = FALSE}
preds <- predict(bert_classifier, dplyr::select(testing_set, word, other_word))
mean(preds$.pred_class == testing_set$target)
# [1] 0.7479572 # YMMV

# We can also try arbitrary input.
predict(
      bert_classifier,
      dplyr::tibble(
        word = c("spacious", "spacious"),
        other_word = c("cramped", "roomy")
      )
    )
# # A tibble: 2 × 1
#   .pred_class
#   <fct>      
# 1 antonym    
# 2 synonym 
```

