---
title: "Fitting a BERT-style Classification Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting a BERT-style Classification Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(tidybert)
```

Here we demonstrate how to easily fit a BERT-based classification model.

## The MNLI Dataset

We will use the [MultiNLI dataset](https://cims.nyu.edu/~sbowman/multinli/) (MNLI) to set up an entailment task. 
For more details about the task and the dataset, see the Textual Entailment vignette in {torchtransformers}.

```{r download-data, options}
# Set up a processor function for {dlr} to load the data.
process_mnli <- function(source_file) {
  dataset_names <- c(
    "train",
    "dev_matched",
    "dev_mismatched",
    "test_matched",
    "test_mismatched"
  )
  # Also make those the names so purrr uses them.
  names(dataset_names) <- dataset_names
  
  mnli_tibbles <- purrr::map(
    dataset_names,
    function(this_dataset) {
      # We specify column types to make sure things come in as we expect.
      column_spec <- dplyr::case_when(
        stringr::str_starts(this_dataset, "dev_") ~ "iicccccccccccccc",
        stringr::str_starts(this_dataset, "test_") ~ "iiiccccccc",
        TRUE ~ "iicccccccccc"
      )
      raw_tibble <- readr::read_tsv(
        unz(source_file, fs::path("MNLI", this_dataset, ext = "tsv")),
        col_types = column_spec,
        # There are a couple lines that screw up if we include a quote
        # character.
        quote = ""
      )
      # If there are labels, standardize them, to make sure the factor levels
      # are always the same.
      if ("gold_label" %in% colnames(raw_tibble)) {
        raw_tibble$gold_label <- factor(
          raw_tibble$gold_label,
          levels = c("entailment", "neutral", "contradiction")
        )
      }
      return(
        dplyr::select(
          raw_tibble,
          -index,
          -promptID,
          -pairID,
          -dplyr::ends_with("_parse"),
          -dplyr::starts_with("label")
        )
      )
    }
  )
  
  return(mnli_tibbles)
}

# By default downloading large files often fails. Increase the timeout.
old_timeout <- options(timeout = 1000)

data_url <- "https://dl.fbaipublicfiles.com/glue/data/MNLI.zip"

mnli_tibbles <- dlr::read_or_cache(
  source_path = data_url,
  appname = "tidybert",
  process_f = process_mnli
)

# Restore the timeout.
options(old_timeout)
```

## Fit the Model

We will use the `train` dataset to fit a simple classification model.
We'll fit small models by various methods, using a subset of the data. 
If you were really training a model, we recommend using much more than 150 rows, and more than a single epoch.
We'll reset seeds each time to confirm that the models are identical.

```{r fit}
train_head <- head(mnli_tibbles$train, 150)

set.seed(1234)
torch::torch_manual_seed(1234)
entailment_fit_formula <- bert_classification(
  gold_label ~ sentence1 + sentence2, 
  data = train_head,
  bert_type = "bert_tiny_uncased",
  n_tokens = 128,
  epochs = 1
)

set.seed(1234)
torch::torch_manual_seed(1234)
entailment_fit_dataframe <- bert_classification(
  x = dplyr::select(train_head, sentence1, sentence2),
  y = train_head$gold_label,
  bert_type = "bert_tiny_uncased",
  n_tokens = 128,
  epochs = 1
)

set.seed(1234)
torch::torch_manual_seed(1234)
entailment_fit_matrix <- bert_classification(
  x = as.matrix(dplyr::select(train_head, sentence1, sentence2)),
  y = train_head$gold_label,
  bert_type = "bert_tiny_uncased",
  n_tokens = 128,
  epochs = 1
)
```

Regardless of your preferred format for providing the data, the resulting models are the same.

## Make Predictions

Pass in new data to make predictions.
The predictions can be in the form of the predicted label ("class", default), or the class probabilities ("prob").

We predict with each method using each of the fitted models to verify that they are the same.

```{r predict}
entailment_predictions_formula <- predict(
  entailment_fit_formula,
  new_data = head(mnli_tibbles$dev_matched)
)

entailment_probs_formula <- predict(
  entailment_fit_formula,
  new_data = head(mnli_tibbles$dev_matched),
  type = "prob"
)
entailment_probs_dataframe <- predict(
  entailment_fit_dataframe,
  new_data = head(mnli_tibbles$dev_matched),
  type = "prob"
)
entailment_probs_matrix <- predict(
  entailment_fit_matrix,
  new_data = head(mnli_tibbles$dev_matched),
  type = "prob"
)

identical(entailment_probs_formula, entailment_probs_dataframe)
identical(entailment_probs_formula, entailment_probs_matrix)
```
